Baseline centering problem

### 2019-01-14 paper/tangotd.tex

Building a universal policy. One may make the policy depend on a goal
representation $\omega$ as an additional input. One may simultaneously
compute $Z(s_1,s_2;\omega)$, namely the $Z$ function when following the
policy with goal $\omega$, and train at the same time a map $s_2\mapsto \omega=\Omega(s_2)$ and
policy parameters $\pi(a|s_1,\omega)$ such that following
$\pi(|\Omega(s_2))$ indeed goes towards $s_2$, namely, the policy
gradient for every $s_2$ is
\begin{equation}
\partial \pi(a|s,\Omega(s_2))\E_{s'|s,a} \E_{s_2\sim \rho} \,
(z(s',s_2;\Omega(s_2))-z(s,s_2;\Omega(s_2)))
\end{equation}
This equation is used to train the policy parameters; $\Omega(s_2)$ is
trained so that $\omega=\Omega(s_2)$ indeed maximizes $z(s,s_2,\omega)$
(because the best way to reach $s_2$ is to select the $\omega$ with
maximal $z(s,s_2,\omega)$ at each step; TODO: from which $s$? TODO: here
I'm confusing constant action plans and time-variable action plans).
[TODO check this is indeed the policy gradient for the time spent at
$s_2$ if following policy $\pi(\Omega(s_2))$]. Advantage: diversity of
policies should appear by itself, no need to manually force options to be
distinct. Depends on a chosen distribution of goals $s_2$; we could learn
to go to uniformly distributed states, or create a goal at points where
we observe a high reward value (so learning to go to all states with a
reward). Just fixing $\Omega(s)\deq s$ once and for all might work, in
principle, but may be ambitious (or use a summary/auto-encoded value of
$s$). TODO check:
this is with a reward a Dirac at $s_2$, check there's no technical
problem wrt the reference distribution $\rho(s_2)$, as we use $z$ not $Z$
in the policy gradient and $z$ is the density of $Z$ wrt $\rho$ but I
think it's correct. Actually, for goal representations, a goal is not
necessarily a state itself, should be a feature of states, or some
representation of a reward function, so we should figure out the way to
give a meaning to $\Omega(R)$ for an arbitrary reward $R$. The
representation of $Z$ as $Z=\transp{\green}f$ does that automatically,
thanks to linearity wrt $R$. This could even help with exploration:
setting temporary goals towards already-seen past states solves the "pure
random walk" problem by introducing temporary momentum; one can even
increase temperature a bit on $\omega$, by using slightly larger/more
variant values of $\omega$ as the goal.

@@

\paragraph{Early version of the above: Learning to navigate.} Define a function $V(s,s')$ that
represents the time spent in $s'$ if starting at $s$ and acting optimally
towards $s'$ (just as if the reward was $\1_{s'}$). It is defined by an iterated expectimax to reach $s'$, and
satisfies the Bellman equation
\begin{equation}
V(s,s')=\1_{s=s'}+\gamma \max_a \E_{s_1|s,a} V(s_1,s')
\end{equation}
and it also satisfies
\begin{equation}
V(s,s')=\1_{s=s'}+\gamma \sum_{s_{t-1}} V(s,s_{t-1})\max_a
\E_{s_t|s_{t-1},a}\1_{s_t=s'}
\end{equation}
although it's not clear to me how to exploit the last expression. 
\footnote{On-policy it works, I think, because on-policy we can sample
reverse transitions. But I think in that case $V(s,s')$ has to be defined
as a density wrt $\mu(s')$.\\Should also work if we maintain a low-rank
estimate of $V$ as a sum of Diracs over a few $s'$.}
But I
think we have to work from both ends if we want to represent
compositionality of actions. Indeed with the first equation, different values
of $s'$ are just completely separate Bellman problems; the generalization
abilities of the function approximator for $V$ may help, but only so
much. Each time we observe a transition $s\to s'$, we should be able to
update the probabilities to reach $s'$ based on the probabilities to
reach $s$.

@@

>>>>

Building a universal policy. One may make the policy depend on a goal
representation $\omega$ as an additional input. One may simultaneously
compute $Z(s_1,s_2;\omega)$, namely the $Z$ function when following the
policy with goal $\omega$, and train at the same time a map $s_2\mapsto \omega=\Omega(s_2)$ and
policy parameters $\pi(a|s_1,\omega)$ such that following
$\pi(|\Omega(s_2))$ indeed goes towards $s_2$, namely, the policy
gradient for every $s_2$ is
\begin{equation}
\partial \pi(a|s,\Omega(s_2))\E_{s'|s,a} \E_{s_2\sim \rho} \,
(z(s',s_2;\Omega(s_2))-z(s,s_2;\Omega(s_2)))
\end{equation}
This equation is used to train the policy parameters; $\Omega(s_2)$ is
trained so that $\omega=\Omega(s_2)$ indeed maximizes $z(s,s_2;\omega)$
(because the best way to reach $s_2$ is to select the $\omega$ with
maximal $z(s,s_2;\omega)$ at each step; TODO: from which $s$? TODO: here
I'm confusing constant action plans and time-variable action plans).
[TODO check this is indeed the policy gradient for the time spent at
$s_2$ if following policy $\pi(\Omega(s_2))$]. Advantage: diversity of
policies should appear by itself, no need to manually force options to be
distinct. Depends on a chosen distribution of goals $s_2$; we could learn
to go to uniformly distributed states, or create a goal at points where
we observe a high reward value (so learning to go to all states with a
reward). Just fixing $\Omega(s)\deq s$ once and for all might work, in
principle, but may be ambitious (or use a summary/auto-encoded value of
$s$). TODO check:
this is with a reward a Dirac at $s_2$, check there's no technical
problem wrt the reference distribution $\rho(s_2)$, as we use $z$ not $Z$
in the policy gradient and $z$ is the density of $Z$ wrt $\rho$ but I
think it's correct. Actually, for goal representations, a goal is not
necessarily a state itself, should be a feature of states, or some
representation of a reward function, so we should figure out the way to
give a meaning to $\Omega(R)$ for an arbitrary reward $R$. The
representation of $Z$ as $Z=\transp{\green}f$ does that automatically,
thanks to linearity wrt $R$. This could even help with exploration:
setting temporary goals towards already-seen past states solves the "pure
random walk" problem by introducing temporary momentum; one can even
increase temperature a bit on $\omega$, by using slightly larger/more
variant values of $\omega$ as the goal.

@@

\paragraph{Early version of the above: Learning to navigate.} Define a function $V(s,s')$ that
represents the time spent in $s'$ if starting at $s$ and acting optimally
towards $s'$ (just as if the reward was $\1_{s'}$). It is defined by an iterated expectimax to reach $s'$, and
satisfies the Bellman equation
\begin{equation}
V(s,s')=\1_{s=s'}+\gamma \max_a \E_{s_1|s,a} V(s_1,s')
\end{equation}
and it also satisfies
\begin{equation}
\label{eq:optrightbellman}
V(s,s')=\1_{s=s'}+\gamma \sum_{s_{t-1}} V(s,s_{t-1})\max_a
\E_{s_t|s_{t-1},a}\1_{s_t=s'}
\end{equation}
although it's not clear to me how to exploit the last expression. 
\footnote{On-policy it works, I think, because on-policy we can sample
reverse transitions. But I think in that case $V(s,s')$ has to be defined
as a density wrt $\mu(s')$.\\Should also work if we maintain a low-rank
estimate of $V$ as a sum of Diracs over a few $s'$.}
But I
think we have to work from both ends if we want to represent
compositionality of actions. Indeed with the first equation, different values
of $s'$ are just completely separate Bellman problems; the generalization
abilities of the function approximator for $V$ may help, but only so
much. Each time we observe a transition $s\to s'$, we should be able to
update the probabilities to reach $s'$ based on the probabilities to
reach $s$.

@@

By multiplying \eqref{eq:optrightbellman} by $V$, one finds the
Newton--Bellman equation for the optimal function $V$ seen as a matrix:
\begin{equation}
\label{eq:optrightbellman}
V^2_{s_1s_2}=V_{s_1s_2}+\gamma \sum_{s_3} V_{s_1s_3}\max_a
\E_{s_4|s_3,a} V_{s_4s_2}
\end{equation}
TODO DOUBLE-CHECK
and representing $V$ by its density wrt $\rho$ namely
$V_{s_1s_2}=z(s_1,s_2)\rho(s_2)$ this suggests the Newton method
\begin{equation}
z(s_1,s_2)\gets 2z(s_1,s_2)-\E_{s_3\sim \rho} z(s_1,s_3)z(s_3,s_2)+\gamma
\E_{s_3\sim \rho} z(s_1,s_3)\max_a \E_{s_4\sim P(s_4|s_3,a)} z(s_4,s_2)
\end{equation}
satisfied by the optimal $z$ function.
